---
title: "itr-actg175"
author: "yichen xu"
date: "2024-11-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# basics
1. resource constraints to try
2. category treatment


# interesting research question
* 1. which covariate to choose? how to set V? variable importance? any group-wise difference? 


# question
1. Here do we use stochastic rule?
2. How do the alg determine the suboptimal rule if realistic = True or maximize = False
3. How to output blip functions?

# plan
1. reduce number of V (different combinations)
2. plot the causl estimand with more resource constraints

# check data

```{r}
library(BART)

data(ACTG175)

?ACTG175
```

Y: cd420
W: age, wtkg, hemo, homo, drugs, karnof, oprior, z30, zprior, preanti, race, gender, str2, strat, cd80
A: treat



# fit the basic model

```{r}


library(data.table)
library(sl3)
library(tmle3)
library(tmle3mopttx)


data <- read.csv("actg175.csv")

mean(data$cd420)

node_list <- list(
  W = c("age", "wtkg", "hemo", "homo", "drugs", "karnof", "oprior", "z30", 
        "zprior", "preanti", "race", "gender", "str2", "strat", "cd80"),
  A = "treat",
  Y = "cd420"
)


lrn_mean <- Lrnr_mean$new()
lrn_glm <- Lrnr_glm_fast$new()
lrn_xgboost_50 <- Lrnr_xgboost$new(nrounds = 50)
lrn_hal <- Lrnr_hal9001$new(max_degree = 2, num_knots = rep(5, length(node_list$W)))


Q_learner <- Lrnr_sl$new(
  learners = list(lrn_glm, lrn_mean, lrn_xgboost_50, lrn_hal),
  metalearner = Lrnr_nnls$new()
)


g_learner <- Lrnr_sl$new(
  learners = list(lrn_glm, lrn_mean, lrn_xgboost_50, lrn_hal),
  metalearner = Lrnr_nnls$new()
)


b_learner <- Lrnr_sl$new(
  learners = list(lrn_glm, lrn_mean, lrn_xgboost_50, lrn_hal),
  metalearner = Lrnr_nnls$new()
)


learner_list <- list(Y = Q_learner, A = g_learner, B = b_learner)


tmle_spec <- tmle3_mopttx_blip_revere(
  V = c("age", "wtkg", "hemo", "homo", "drugs", "karnof", "oprior", "z30", 
        "zprior", "preanti", "race", "gender", "str2", "strat", "cd80"), 
  type = "blip1",  
  learners = learner_list,
  maximize = TRUE,
  complex = TRUE,
  realistic = TRUE,
  resource = 1,
  interpret = TRUE
)


fit <- tmle3(tmle_spec, data, node_list, learner_list)


print(fit)



```


# extract rules of basic model

```{r}

table(tmle_spec$return_rule)
zero_indices <- which(tmle_spec$return_rule == 0)
print(zero_indices)

```


# check the positivity assumption

```{r}

library(data.table)
library(sl3)


# Define learners for Super Learner
lrn_mean <- Lrnr_mean$new()
lrn_glm <- Lrnr_glm_fast$new()
lrn_xgboost_50 <- Lrnr_xgboost$new(nrounds = 50)
lrn_hal <- Lrnr_hal9001$new(max_degree = 2, num_knots = rep(5, length(node_list$W)))

# Define the Super Learner for the propensity score model
g_learner <- Lrnr_sl$new(
  learners = list(lrn_glm, lrn_mean, lrn_xgboost_50, lrn_hal),
  metalearner = Lrnr_nnls$new()
)

# Create a task for the propensity score estimation
g_task <- sl3_Task$new(
  data = data,
  covariates = node_list$W,
  outcome = node_list$A,
  outcome_type = "binomial"
)

# Train the Super Learner on the task
g_fit <- g_learner$train(g_task)

# Predict propensity scores P(A = 1 | W)
propensity_scores <- g_fit$predict()

# Compute 1 / P(A = 1 | W)
inv_propensity_scores <- 1 / propensity_scores

# Plot the histogram of 1 / P(A = 1 | W)
hist(inv_propensity_scores, 
     main = "Distribution of 1 / P(A = 1 | W)",
     xlab = "1 / P(A = 1 | W)",
     col = "lightblue", 
     border = "white", 
     breaks = 30)

```







# resource constraint 

```{r}

# resource constraints: 0.9, 0.8, 0.7, 0.6, 0.5
# store fit$summary$lower, fit$summary$upper, fit$init_est, fit$tmle_est


resource_constraints <- c(0.9, 0.8, 0.7, 0.6, 0.5)

results_matrix <- matrix(NA, nrow = length(resource_constraints), ncol = 4)
colnames(results_matrix) <- c("lower", "upper", "init_est", "tmle_est")
rownames(results_matrix) <- resource_constraints

for (i in seq_along(resource_constraints)) {
  resource_constraint <- resource_constraints[i]
  
  tmle_spec <- tmle3_mopttx_blip_revere(
    V = c("age", "wtkg", "hemo", "homo", "drugs", "karnof", "oprior", "z30", 
          "zprior", "preanti", "race", "gender", "str2", "strat", "cd80"), 
    type = "blip1",  
    learners = learner_list,
    maximize = TRUE,
    complex = TRUE,
    realistic = TRUE,
    resource = resource_constraint
  )
  
  fit <- tmle3(tmle_spec, data, node_list, learner_list)
  
  fit_summary <- fit$summary
  results_matrix[i, ] <- c(fit_summary$lower, fit_summary$upper, fit_summary$init_est, fit_summary$tmle_est)
}

results_df <- as.data.frame(results_matrix)
results_df$resource_constraint <- resource_constraints

library(ggplot2)

ggplot(results_df, aes(x = resource_constraint)) +
  geom_point(aes(y = tmle_est, color = "TMLE Estimate"), size = 3) +
  geom_point(aes(y = init_est, color = "Initial Estimate"), size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper, color = "TMLE Estimate"), width = 0.05) +
  scale_x_reverse() +
  labs(title = "TMLE and Initial Estimates with Confidence Intervals Across Resource Constraints",
       x = "Resource Constraint",
       y = "Estimate",
       color = "Legend") +
  theme_minimal()

```



# cd496

```{r}

library(dplyr)

data_clean_na <- data  %>% filter(!is.na(cd496))

mean(data_clean_na$cd496)

node_list_96 <- list(
  W = c("age", "wtkg", "hemo", "homo", "drugs", "karnof", "oprior", "z30", 
        "zprior", "preanti", "race", "gender", "str2", "strat", "cd80"),
  A = "treat",
  Y = "cd496"
)


lrn_mean <- Lrnr_mean$new()
lrn_glm <- Lrnr_glm_fast$new()
lrn_xgboost_50 <- Lrnr_xgboost$new(nrounds = 50)
lrn_hal <- Lrnr_hal9001$new(max_degree = 2, num_knots = rep(5, length(node_list$W)))


Q_learner <- Lrnr_sl$new(
  learners = list(lrn_glm, lrn_mean, lrn_xgboost_50, lrn_hal),
  metalearner = Lrnr_nnls$new()
)


g_learner <- Lrnr_sl$new(
  learners = list(lrn_glm, lrn_mean, lrn_xgboost_50, lrn_hal),
  metalearner = Lrnr_nnls$new()
)


b_learner <- Lrnr_sl$new(
  learners = list(lrn_glm, lrn_mean, lrn_xgboost_50, lrn_hal),
  metalearner = Lrnr_nnls$new()
)


learner_list <- list(Y = Q_learner, A = g_learner, B = b_learner)


tmle_spec_96 <- tmle3_mopttx_blip_revere(
  V = c("age", "wtkg", "hemo", "homo", "drugs", "karnof", "oprior", "z30", 
        "zprior", "preanti", "race", "gender", "str2", "strat", "cd80"), 
  type = "blip1",  
  learners = learner_list,
  maximize = TRUE,
  complex = TRUE,
  realistic = TRUE,
  resource = 1,
  interpret = TRUE
)


fit_96 <- tmle3(tmle_spec_96, data_clean_na, node_list_96, learner_list)


print(fit_96)


table(tmle_spec_96$return_rule)




```

```{r}

library(ggplot2)

# Extract predictions
blip_predictions <- tmle_spec_96$blip_fit$predict()

# Combine into a data frame for ggplot2
plot_data <- data.frame(cd80 = data_clean_na$cd80, blip_predictions = blip_predictions)

# Create scatter plot with smooth curve
ggplot(plot_data, aes(x = cd80, y = blip_predictions)) +
  geom_point(alpha = 0.6) +  # Points
  geom_smooth(method = "loess", se = FALSE, color = "blue") +  # Smooth curve
  labs(
    title = "Blip Predictions vs. cd80",
    x = "cd80",
    y = "Blip Predictions"
  ) +
  theme_minimal()


```


```{r}

hist(tmle_spec_96$blip_fit$predict())

```



```{r}


# resource constraints: 0.9, 0.8, 0.7, 0.6, 0.5
# store fit$summary$lower, fit$summary$upper, fit$init_est, fit$tmle_est


resource_constraints <- c(0.9, 0.8, 0.7, 0.6, 0.5)

results_matrix <- matrix(NA, nrow = length(resource_constraints), ncol = 4)
colnames(results_matrix) <- c("lower", "upper", "init_est", "tmle_est")
rownames(results_matrix) <- resource_constraints

for (i in seq_along(resource_constraints)) {
  resource_constraint <- resource_constraints[i]
  
  tmle_spec <- tmle3_mopttx_blip_revere(
    V = c("age", "wtkg", "hemo", "homo", "drugs", "karnof", "oprior", "z30", 
          "zprior", "preanti", "race", "gender", "str2", "strat", "cd80"), 
    type = "blip1",  
    learners = learner_list,
    maximize = TRUE,
    complex = TRUE,
    realistic = TRUE,
    resource = resource_constraint
  )
  
  fit <- tmle3(tmle_spec_96, data_clean_na, node_list_96, learner_list)
  
  fit_summary <- fit$summary
  results_matrix[i, ] <- c(fit_summary$lower, fit_summary$upper, fit_summary$init_est, fit_summary$tmle_est)
}

results_df <- as.data.frame(results_matrix)
results_df$resource_constraint <- resource_constraints

library(ggplot2)

ggplot(results_df, aes(x = resource_constraint)) +
  geom_point(aes(y = tmle_est, color = "TMLE Estimate"), size = 3) +
  geom_point(aes(y = init_est, color = "Initial Estimate"), size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper, color = "TMLE Estimate"), width = 0.05) +
  scale_x_reverse() +
  labs(title = "TMLE and Initial Estimates with Confidence Intervals Across Resource Constraints",
       x = "Resource Constraint",
       y = "Estimate",
       color = "Legend") +
  theme_minimal()

```


